Modal Abstractions for Virtualizing Memory Addresses

### 2024-06-29 examples.tex
+

As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v}.
Beyond a frame ($P$), the precondition requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
(in \lstinline|rbx|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
The latter uses an additional ghost set added to the per-address-space
invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
that physical memory, and also guarantees the offset is the correct calculation.
Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
\subsubsection{Traversing Page Tables}
As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table. In Figure \ref{fig:getnexttable}, we see how to obtain a physicall address of the next level page-table. 
\begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$   
push   rbp
mov    rsp,rbp
sub    0x20,rsp
mov    rdi,-0x8[rbp]
;;  pte_t *next;

>>>>

As part of the full page table walk verification, we have verified the critical step for a small x86-64 kernel, which is the physical-to-virtual
conversion (often appearing as a macro \texttt{P2V} in C source code). In our small kernel, as in larger kernels, \texttt{P2V} is actually just addition,
but the correctness of this simple instruction is quite subtle, as shown in Figure \ref{fig:p2v}.
Beyond a frame ($P$), the precondition requires the current address space invariant, a check that the addition will not overflow, the physical address to convert
(in \lstinline|rbx|), knowledge that the current address space is valid, and knowledge that the physical address is in a range dedicated
to page table entries (as this physical-to-virtual conversion is only guaranteed for page table memory). 
The latter uses an additional ghost set added to the per-address-space
invariant per Figure \ref{fig:peraspaceinvariant_with_p2v_extension}. The extended invariant owns, for each memory location dedicated to page tables, some fraction of
that physical memory, and also guarantees the offset is the correct calculation.
Subtly, the operation ensures the relevant table entry is readable, but the exact portion of ownership returned must be determined by inspecting the valid bit
of the value in memory --- so full ownership is returned only for unused entries.
This is a simple piece of code whose functionality is critical and whose correctness is highly non-trivial. No prior work engages with this problem.
\subsubsection{Traversing Page Tables}
As we explain in Section \ref{}, a traversal over a page-table tree can be done through the indices into the tables at each level followed by obtaining the address of the next table. In Figure \ref{fig:getnexttable}, we see how to obtain the physicall address of the next level page-table.

\begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
;;void pte_initialize(pte_t **table) {
push   rbp
mov    rsp,rbp
sub    0x10,rsp
mov    rdi,-0x8[rbp]

;;Allocate a full page for 512 8-byte entries
;;pte_t *local = kalloc();
;;clear the entire table. Ensures present (and reserved, etc.) bits are 0
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp]

;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]
;;return;
add    0x10,rsp
pop    rbp
retq   
  \end{lstlisting}
\vspace{-1em}
\caption{ in Figure \ref{fig:getnexttable}.}
\label{fig:pteinitialize}
\end{figure}

\begin{figure}\footnotesize
 \begin{lstlisting}[mathescape]
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
;; pte_t *pte_get_next_table_succ(pte_t *entry) {
push   rbp
mov    rsp,rbp
sub    0x20,rsp
mov    rdi,-0x8[rbp]
;;  pte_t *next;
